---
title: "Airline Data"
author: "tygeremily"
date: "01/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(RSQLite)
library(ggplot2)

delay.con <- dbConnect(RSQLite::SQLite(),
                       dbname = "~/AirlineDelay.sqlite3")
```

1. What was the most popular airline in 2000?
2. How many flights did that airline record in 2000?
```{r}
popair = dbGetQuery(delay.con,            
               "SELECT COUNT(*), UniqueCarrier 
               FROM AirlineDelay 
               WHERE Year=2000 
               GROUP BY UniqueCarrier 
               ORDER BY COUNT(*) DESC LIMIT 10 ")
# View(popair)
```

3. Since 2005, which airline has undertaken the longest flights (time) on average?
```{r}
longflight = dbGetQuery(delay.con, 
              "SELECT AVG(ActualElapsedTime) AS Elapsedtime, UniqueCarrier           
              FROM AirlineDelay           
              WHERE (ActualElapsedTime IS NOT 'NA') AND Year >= 2005           
              GROUP BY UniqueCarrier            
              ORDER BY Elapsedtime DESC LIMIT 10")
# View(longflight)
```

4. How many flights used a carrier with an E in the name? (e.g. XE)
```{r}
Ecarriers = dbGetQuery(delay.con,            
               "SELECT COUNT(*), UniqueCarrier            
           FROM AirlineDelay            
           WHERE UniqueCarrier LIKE '%E%'            
           GROUP BY UniqueCarrier ")
# View(Ecarriers)
```

5. For each of the 22 years in the dataset, how many of the tail codes are listed as NA? 
Paste the table of values.
```{r}
tailNAs = dbGetQuery(delay.con, 
                      "SELECT COUNT(*), Year            
                      FROM AirlineDelay            
                      WHERE (TailNum IS 'NA') AND Year >= 2000
                      GROUP BY Year")
#View(tailNAs)
```

6. Which of the following Origin->Destination combinations have the largest (average) Arrival delays?
```{r}
largedelay = dbGetQuery(delay.con,            
                         "SELECT COUNT(*), Origin || Dest as combined, 
                         AVG(ArrDelay) as Avg_ArrDelay           
                         FROM AirlineDelay            
                         GROUP BY combined          
                         ORDER BY Avg_ArrDelay DESC")
# View(largedelay)
```

7. Are flights being given more (average) time to reach their destination for later years? Use ggplot to display the trend for each year in the dataset, and upload the image.
```{r}
Flighttimebyyear = dbGetQuery(delay.con,            
                              "SELECT Year, AVG(CRSElapsedTime) as Time         
                              FROM AirlineDelay            
                              GROUP BY Year")
# View(Flighttimebyyear)
Flighttimebyyear = Flighttimebyyear[1:22,]

timeyearplot = ggplot(Flighttimebyyear, aes(x = Year, y = Time)) +
  geom_point() 
timeyearplot

ggsave("timeyearplot.png", plot = timeyearplot)
```

8. As a possible explanation, are the average flight distances getting longer every year? Use ggplot to display the trend for each year in the dataset, and upload the image.
```{r}
Flightdistancebyyear = dbGetQuery(delay.con,            
                              "SELECT Year, AVG(Distance) as Distance          
                              FROM AirlineDelay            
                              GROUP BY Year")
# View(Flightdistancebyyear)
Flightdistancebyyear = Flightdistancebyyear[1:22,]

distyearplot = ggplot(Flightdistancebyyear, aes(x = Year, y = Distance)) +
  geom_point() 
distyearplot

ggsave("distyearplot.png", plot = distyearplot)
```

9. Compute the correlation (aka correlation coefficient) between the two quantities in the question above? I.e. the correlation between 1) The average flight distance for each year 2) The average scheduled time per flight for each year
```{r}
temp1 = list(Flighttimebyyear, Flightdistancebyyear)
timedist <- Reduce(function(d1,d2) merge(d1,d2, by = c("Year"), all.x = TRUE), temp1)
# View(timedist)

corr = cor.test(timedist$Year, timedist$Distance)
corr
```

10. Find the total number of delayed arrivals for each year (ArrDelay>0). Find the total number of delayed departures for each year (DepDelay>0). Fit a linear model between the two quantities and give the equation of the line (intercept and slope)
```{r}
delArr = dbGetQuery(delay.con,            
                    "SELECT COUNT(*) as arrival, Year           
                    FROM AirlineDelay
                    WHERE ArrDelay > 0
                    GROUP BY Year")
# View(delArr)
delArr = delArr[1:22,]

delDep = dbGetQuery(delay.con,            
                    "SELECT COUNT(*) as departure, Year           
                    FROM AirlineDelay 
                    WHERE DepDelay > 0 
                    GROUP BY Year")
# View(delDep)
delDep = delDep[1:22,]

temp2 = list(delArr, delDep)
delays <- Reduce(function(d1,d2) merge(d1,d2, by = c("Year"), all.x = TRUE), temp2)
# View(delays)

model = lm(Year ~ arrival + departure, data = delays)
model
```
